### 2015-12-07 mirror-descent.tex
Fixed typos in mirror descent section.

The regret bound for \textsc{Scale-Free Mirror Descent}
in the Corollary~\ref{corollary:regret-scale-free-mirror-descent}
depends on $\sup_{u,v \in K} \Breg_f(u,v)$. In contrast,
the regret bounds for \textsc{AdaFTRL} and \textsc{SOLO FTRL}
in Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
depend on $\sup_{u \in K} f(u)$.
Similarly, regret bound in Theorem~\ref{theorem:regret-scale-free-mirror-descent} for
\textsc{Scale-Free Mirror Descent} depends on $\sup_{v \in K} \Breg_R(u,v)$ and
regret bounds in Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to show that
$$
\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v)
$$
provided that at the minimizer $v^* = \argmin_{v \in K} R(v)$ both $R(v^*)$ and
$\grad R(v^*)$ are zero. Indeed, in that case, $R(u) = \Breg_R(u,v^*) \le
\sup_{v \in K} \Breg_R(u,v)$.

The assumption $R(v^*) = 0$ and $\grad R(v^*) = 0$
are easy to achieve by adding an affine function to the regularizer:
$$
R'(u) = R(u) - \langle \grad R(v^*), u - v^* \rangle - R(v^*) \; .
$$
The regularizer $R'$ has the same parameter of strong convexity as $R$, the
associated Bregman divergences $D_{R'}$ and $D_{R}$ are equal, $R$ and $R'$
have the same minimizer $v^*$, and $R'(v^*)$ and $\grad R'(v^*)$ are both zero.

Thus, ignoring constant factors, regret bound for \textsc{Scale-Free Mirror
Descent} is worse than the regret bounds for \textsc{AdaFTRL} and \textsc{SOLO
FTRL}.  In fact, there are examples where $R(u)$ is finite whereas $\sup_{v \in
K} \Breg_R(u,v)$ is infinite. We mention two examples. The first example is $R(w) =
\frac{1}{2}\|w\|_2^2$ defined on the whole space $V$, where for any $u \in V$,
$R(u)$ is a finite value but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in V}
\frac{1}{2}\|u-v\|_2^2 = +\infty$. The second example is the negative entropy
regularizer $R(w) = \ln d  + \sum_{j=1}^d w_j \ln w_j$ defined on the
$d$-dimensional probability simplex $K = \{ w \in \R^d ~:~ w_j \ge 0,
\sum_{j=1}^d w_j = 1 \}$, where for any $u \in K$, $R(u)$ is finite and in fact
lies in the interval $[-\ln d, 0]$ but $\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in K}
\sum_{j=1}^d u_j \ln(u_j/v_j) = + \infty$. We return to these examples in the
following subsection.

@@

The second counter-example is stated as Theorem~\ref{theorem:second-counter-example}
below.  The decision set is the $d$-dimensional probability simplex $K = \{ w
\in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ and the regularizer is the
negative entropy $R(w) = \sum_{j=1}^d w_j \ln w_j$.  Negative entropy is
$1$-strongly convex with respect to $\|\cdot\|_1$ and the dual norm of
$\|\cdot\|_1$ is \mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman
divergence is the Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{j=1}^d u_j
\ln(u_j/v_j)$.  Note that despite that negative entropy is upper- and
lower-bounded, Kullback-Leibler divergence is not.

\begin{theorem}[Second Counter-Example]
\label{theorem:second-counter-example}
Let $d \ge 2$ and let $K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}
\subseteq V$ be the $d$-dimensional probability simplex.  For any $T \ge 120$
there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$
such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots = \|\ell_T\|_\infty =
1$ and \textsc{Scale-Free Mirror Descent} with regularizer $R(w) = \sum_{j=1}^d
w_j \ln w_j$ satisfies
$$
\Regret_T \ge \frac{T}{6} \; .
$$
\end{theorem}

>>>>

The regret bound for \textsc{Scale-Free Mirror Descent}
in the Corollary~\ref{corollary:regret-scale-free-mirror-descent}
depends on $\sup_{u,v \in K} \Breg_f(u,v)$. In contrast,
the regret bounds for \textsc{AdaFTRL} and \textsc{SOLO FTRL}
in Corollaries~\ref{corollary:ada-ftrl-regret-bound}~and~\ref{corollary:regret-solo-ftrl-bounded-set}
depend on $\sup_{u \in K} f(u)$.
Similarly, regret bound in Theorem~\ref{theorem:regret-scale-free-mirror-descent} for
\textsc{Scale-Free Mirror Descent} depends on $\sup_{v \in K} \Breg_R(u,v)$ and
regret bounds in Theorems~\ref{theorem:ada-ftrl-regret-bound}~and~\ref{theorem:regret-solo-ftrl}
for \textsc{AdaFTRL} and \textsc{SOLO FTRL} depend on $R(u)$. It is not hard to show that
\begin{equation}
\label{equation:regularizer-vs-divergence}
\forall u \in K \qquad R(u) \le \sup_{v \in K} \Breg_R(u,v)
\end{equation}
provided that at the minimizer $v^* = \argmin_{v \in K} R(v)$ both $R(v^*)$ and
$\grad R(v^*)$ are zero. Indeed, in that case, $R(u) = \Breg_R(u,v^*) \le
\sup_{v \in K} \Breg_R(u,v)$.

The assumption $R(v^*) = 0$ and $\grad R(v^*) = 0$
are easy to achieve by adding an affine function to the regularizer:
$$
R'(u) = R(u) - \langle \grad R(v^*), u - v^* \rangle - R(v^*) \; .
$$
The regularizer $R'$ has the same parameter of strong convexity as $R$, the
associated Bregman divergences $\Breg_{R'}$ and $\Breg_{R}$ are equal, $R'$ and
$R$ have the same minimizer $v^*$, and $\grad R'(v^*)$ and $\grad R'(v^*)$ are
both zero.

Thus, \eqref{equation:regularizer-vs-divergence} implies that ignoring constant
factors, regret bound for \textsc{Scale-Free Mirror Descent} is \emph{not}
better and possibly worse than the regret bounds for \textsc{AdaFTRL} and
\textsc{SOLO FTRL}.  In fact, it is not hard to come up with examples where
$R(u)$ is finite whereas $\sup_{v \in K} \Breg_R(u,v)$ is infinite. We mention
two such examples. The first example is $R(w) = \frac{1}{2}\|w\|_2^2$ defined
on the whole space $V$, where for any $u \in V$, $R(u)$ is a finite value but
$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in V} \frac{1}{2}\|u-v\|_2^2 =
+\infty$. The second example is the negative entropy regularizer $R(w) = \ln d
+ \sum_{j=1}^d w_j \ln w_j$ defined on the $d$-dimensional probability simplex
$K = \{ w \in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$, where for any $u
\in K$, $R(u)$ is finite and in fact lies in the interval $[-\ln d, 0]$ but
$\sup_{v \in K} \Breg_R(u,v) = \sup_{v \in K} \sum_{j=1}^d u_j \ln(u_j/v_j) = +
\infty$. We return to these examples in the following subsection.

@@

The second counter-example is stated as Theorem~\ref{theorem:second-counter-example}
below.  The decision set is the $d$-dimensional probability simplex $K = \{ w
\in \R^d ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$ and the regularizer is the
negative entropy $R(w) = \sum_{j=1}^d w_j \ln w_j$.  Negative entropy is
$1$-strongly convex with respect to $\|\cdot\|_1$ and the dual norm of
$\|\cdot\|_1$ is \mbox{$\|\cdot\|_\infty$}.  The corresponding Bregman
divergence is the Kullback-Leibler divergence $\Breg_R(u,v) = \sum_{j=1}^d u_j
\ln(u_j/v_j)$.  Note that despite that negative entropy is upper- and
lower-bounded, Kullback-Leibler divergence is unbounded from above.

\begin{theorem}[Second Counter-Example]
\label{theorem:second-counter-example}
Let $d \ge 2$, let $V = \R^d$, and let $K = \{ w \in V ~:~ w_j \ge 0, \sum_{j=1}^d w_j = 1 \}$
be the $d$-dimensional probability simplex.  For any $T \ge 120$
there exists a sequence of loss vectors $\ell_1, \ell_2, \dots, \ell_T \in V^*$
such that $\|\ell_1\|_\infty = \|\ell_2\|_\infty = \dots = \|\ell_T\|_\infty =
1$ and \textsc{Scale-Free Mirror Descent} with regularizer $R(w) = \sum_{j=1}^d
w_j \ln w_j$ satisfies
$$
\Regret_T \ge \frac{T}{6} \; .
$$
\end{theorem}
